Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change

### 2020-11-25 Scripts/Manuscript.Rmd

---
output:
  bookdown::pdf_book:
    toc : no
    keep_tex: no
    number_sections: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/ToE_Ref.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{float} # Use the 'float' package
- \floatplacement{figure}{H} # Make every figure with caption = h
---

<!-- --- -->
<!-- title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change -->
<!-- subtitle: Challenges to transboundary fisheries under climate change -->
<!-- author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^' -->
<!-- date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern.' -->
<!-- output: -->
<!--   # word_document:  -->
<!--     # reference_docx: ../References/template.docx -->
<!--   pdf_document: default -->
<!-- bibliography: ../References/ToE_Ref.bib -->
<!-- csl: ../References/science.csl -->
<!-- header-includes: -->
<!-- - \usepackage{amsmath} -->
<!-- - \usepackage{float} # Use the 'float' package -->
<!-- - \floatplacement{figure}{H} # Make every figure with caption = h -->
<!-- --- -->

@@

Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn].  Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]. ). As a result of climate change, many marine species have changed their distribution towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk]. Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt] impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including managing fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].

The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl]. As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions including both the high seas and EEZs [@Munro:2004th]. While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl]. A recent study estimates that, globally, there are 633 transboundary fish species, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].

@@

Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide. We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods). We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study). Our results suggest that many transboundary stocks have already ‘emerged’ from historical natural variability and overpassed their share threat point in most neighboring EEZs, posing a threat to the sustainability of shared stocks and the resilience of international treaties.

@@

The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 \pm 26 years (Figure \@ref(fig:toeFigure1)). In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100. Fish stocks with transboundary index that first emerged from their historical variability in each EEZ was projected to be as early as 2006 and the last one in 2100. About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g. Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010). The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020. After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.7% per year. By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure Figure \@ref(fig:toeFigure1)). Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020. Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century and are not expected to stop there (Figure Figure \@ref(fig:toeFigure1)). We based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift). Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g. the top 90%) of the stock (Figure Figure \@ref(fig:toeFigures·3).

@@

The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A). Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1)) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B). In contrast, EEZs located in template regions like northern Europe and eastern Asia will have a significantly later time of emergence (p < 0.05; see Table \@ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058). However, notable exceptions of tropical EEZs with a later average time of emergence across stocks include Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070). Moreover, despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) will have relative early average time of emergence (Figure \@ref(fig:toeFigure2)A). Only in a few cases, the transboundary index of a stock will not emerge between 2006 and 2100. In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*). In the case of Brazil, we included four transboundary stocks, all shared with Uruguay [@PalaciosAbrantes:2020]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*). In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution \@ref(fig:toeFigure2)A)

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of stocks’ transboundary index. A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks. Aqua color represents EEZs with no time of emergence between 2006 and 2100. B) Time of emergence of the transboundary index according to the United Nations sub-regions. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent points outside of 1.5*IQR."}

@@

On average, stocks that will see their transboundary index emerge by 2100 represent 27% ± 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010. However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs. In some cases like El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries  in their respective EEZs between 2004 and 2014. On the contrary, stocks with emerging transboundary index from the Marshal Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A). In some cases, a few emerging stocks represent large proportion of revenue from transboundary stocks such as Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively). On the contrary, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the participation of such stocks on fishing revenue is less than 20%.

@@

We set an arbitrary threshold of 1 s.d. (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods). We tested the sensitivity of our methods by considering a larger threshold for the time of emergence results (i.e., 2 s.d for a 95% confidence). The global average time of emergence of the transboundary index considering a 2 s.d. threshold was 2036 \pm 28, a difference of about 7 years relative to the 1 s.d. threshold (Figure \@ref(fig:toeFigureS2)). In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric, that is, 40% (n = 1626) fewer emerging stocks than using a 1.s.d threshold.

@@

```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association. Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Methods). The number of species included in this analysis for each habitat type is noted in parenthesis. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers outside of 1.5 times the interquartile range."}

@@

We have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)). By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% \pm 15% (Figure \@ref(fig:toeFigure5)A). The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)). However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high. For example, Guatemala shares Panulirus gracilis stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century. Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)). Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A). These EEZs will see one third (33% \pm 21%) of their transboundary stocks changing the share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A – Land). Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Viet Nam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.

@@

Our results offer no clear global pattern in the change of stock share ratio (Figure \@ref(fig:toeFigure5)). In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, stock share ratio is expected to follow a poleward shift. However, central America Pacific and West African coasts are expected to follow an equatorial direction. Moreover, some regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no stablished pattern, this is the case of the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).

@@

Our projections in the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century corroborate with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk]. For example, in the early 2000s, Humboldt squid (Dosidicus gigas) substantially expanded its geographic range poleward, reaching the coast of Washington state (US). A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs]. In the northeast Atlantic, Atlantic mackerel (Scomber scombrus) is multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC). However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008 without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij]. Such changes in quota resulted in disputes between Iceland and Faroes Islands, and the NEAFC member states [@Spijkers:2017ij]. As transboundary stocks continue to emerge in the incoming years, imal-adapted fisheries management and international disputes like that of A. mackerel are likely to increase in frequency [@Pinksy2018]

The projected range shifts of large number of transboundary stocks emerging since the early 2000s can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy]. Marine organisms have a thermal tolerance that limits the temperature range they could live in [@IPCC:2019tn]. This thermal tolerance explains the sensitivity of the animal to warming and is a main driver of species’ distributional change [@IPCC:2019tn]. Oxygen plays a similar role to that of temperature [Clarke:2020ep], especially on tuna species that require large amounts of dissolve oxygen for they metabolism [@IPCC2014]. It has been estimated that since the early 2000s, increase in sea surface temperature (SST) and decrease in oxygen content have emerged in 79% and 23% of the global ocean, respectively, and by 2080, SST is projected to emerge in 90% of the global ocean while O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq]. Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario respectively [@IPCC:2019tn]. These oceanographic changes are important drivers of change in species distributions as represented in the DBEM [@Cheung:2016jd; @Cheung:0ur].

@@

The lack of Arctic stocks with emerging transboundary index is likely a combination of the few species currently fished in this region. Historically, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg]. However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@REF]. As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh]. Moreover, projections suggest a larger increment on transboundary stocks in the incoming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further North [@Morley:2018fn; @Frainer:2017ji]. For Brazil and Uruguay, the current and future distribution of the transboundary stocks in combination with the particular characteristics of these EEZs could be influencing the results. First, both Argentine menhaden and Argentine shortfin squid have distributions with larger biomass concentration further south, with Brazil serving as the northern limit of the distributional range [@AquaMaps:2019ui; @AquaMaps:2019wt]. Second, Brazil’s EEZ is substantially larger than Uruguay’s resulting in a long distance between Brazil’s and the stocks’ centroids. Third, both stocks are expected to shift further south in the incoming years [@AquaMaps:2019ui; @AquaMaps:2019w]. Thus, the further southern shift of Argentine menhaden and shortfin squid in combination with the reduced latitudinal gradient of Uruguay’s EEZ results in a small effect in the already long latitudinal distance between Brazil’s EEZ and the stocks’ centroids. In the case of the two shrimp species, their destitution range spams along the cost of Brazil and is not expected to have a strong shift in the incoming years, also resulting in a stable transboundary index [@AquaMaps:2019wn; @AquaMaps:2019wj].

>>>>

---
title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
subtitle: Challenges to transboundary fisheries under climate change
author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^'
date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern.'
output:
  # word_document:
    # reference_docx: ../References/template.docx
  pdf_document: default
bibliography: ../References/ToE_Ref.bib
csl: ../References/science.csl
header-includes:
- \usepackage{amsmath}
- \usepackage{float} # Use the 'float' package
- \floatplacement{figure}{H} # Make every figure with caption = h
---

@@

Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn]. Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]). As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk]. Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including the management of fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].

The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl]. As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary stocks, those that cross neighboring EEZs; (*ii*) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs [@Munro:2004th]. While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl]. A recent study estimates that, globally, there are 633 transboundary fish species, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].

@@

Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide. We employ a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods). We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study). Our results suggest that many transboundary stocks have already ‘emerged’ from historical natural variability and overpassed their share threat point in most neighboring EEZs, posing a threat to the sustainability of shared stocks and the resilience of international treaties.

@@

The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 \pm 26 years (Figure \@ref(fig:toeFigure1)). In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100. Fish stocks with transboundary index that first emerged from their historical variability in each EEZ was projected to be as early as 2006 and the last one in 2100. About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g. Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010). The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020. After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.7% per year. By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)). Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020. Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure Figure \@ref(fig:toeFigure1)). We based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift). Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g. the top 90%) of the stock (Figure \@ref(fig:toeFigureS3).

@@

The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A). Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1)) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B). In contrast, EEZs located in template regions like northern Europe and eastern Asia will have a significantly later time of emergence (p < 0.05; see Table \@ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058). However, notable exceptions of tropical EEZs with a later average time of emergence across stocks include Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070). Moreover, despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) will have relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A). Only in a few cases, the transboundary index of a stock will not emerge between 2006 and 2100. In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*). In the case of Brazil, we included four transboundary stocks, all shared with Uruguay [@PalaciosAbrantes:2020tv]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*). In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A)

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of stocks’ transboundary index. A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks. Aqua color represents EEZs with no time of emergence between 2006 and 2100. B) Time of emergence of the transboundary index according to the United Nations sub-regions. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}

@@

On average, stocks that will see their transboundary index emerge by 2100 represent 27% \pm 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010. However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs. In some cases e.g., El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries  in their respective EEZs between 2004 and 2014. On the contrary, stocks with emerging transboundary index from the Marshal Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A). In addition, some EEZs are projectede to have a few emerging stocks that represent large proportion of revenue from transboundary stocks e.g., in Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively). Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks on fishing revenue is less than 20%.

@@

We set an arbitrary threshold of 1 s.d. (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods). We tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence). The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 \pm 28, a difference of about 7 years relative to the 1 s.d. threshold (Figure \@ref(fig:toeFigureS2)). In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric, that is, 40% (n = 1626) fewer emerging stocks than using a 1.s.d threshold.

@@

```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association. Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Methods). The number of species included in this analysis for each habitat type is noted in parenthesis. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}

@@

We have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)). By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% \pm 15% (Figure \@ref(fig:toeFigure5)A). The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)). However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high. For example, Guatemala shares Panulirus gracilis stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century. Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)). Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A). These EEZs will see one third (33% \pm 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A). Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.

@@

Changes of stock share ratio are largely related to regional changes in biogeography and the geometry of the EEZs (Figure \@ref(fig:toeFigure5)). In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change. However, central America Pacific and West African coasts are expected to follow an equatorial direction as range shifts of the transboundary species are following local oceanographic gradients with relatively lower temperature waters in lower latitude [@Clarke:2020ep]. Moreover, some regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no stablished pattern, this is the case of the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).

@@

Our projections in the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century corroborate with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk]. For example, in the early 2000s, Humboldt squid (Dosidicus gigas) substantially expanded its geographic range poleward, reaching the coast of Washington state (US). A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs]. In the northeast Atlantic, the Atlantic mackerel (Scomber scombrus) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC). However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008 without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij]. Such changes in quota resulted in disputes between Iceland and Faroes Islands, and the NEAFC member states [@Spijkers:2017ij]. As transboundary stocks continue to emerge in the incoming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb]

The projected range shifts of large number of transboundary stocks emerging since the early 2000s can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy]. Marine organisms have a thermal tolerance that limits the temperature range they could live in [@IPCC:2019tn]. This thermal tolerance explains the sensitivity of the animal to warming and is a main driver of species’ distributional change [@IPCC:2019tn]. Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially on tuna species that require large amounts of dissolve oxygen for their metabolism [@IPCC:2014bz]. It has been estimated that since the early 2000s, increase in sea surface temperature (SST) and decrease in oxygen content have emerged in 79% and 23% of the global ocean, respectively, and by 2080, SST is projected to emerge in 90% of the global ocean while O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq]. Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario, respectively [@IPCC:2019tn]. These oceanographic changes are important drivers of change in species distributions as represented in the DBEM [@Cheung:2016jd; @Cheung:0ur].

@@

The presentce of EEZs lacking stocks with emerging transboundary index is a combination of EEZ shape, climate projections, and a reduced number of transbounudary stocks. For example, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg]. However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@REF]. As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh]. Moreover, projections suggest a larger increment on transboundary stocks in the incoming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further North [@Morley:2018fn; @Frainer:2017ji]. In the Case of Brazil, in addition to having few transbounudary stocks (only 4 shared with Uruguay in this study), the large diffrence in the EEZs lenghts could also be influencing the results.

<!-- For Brazil and Uruguay, the current and future distribution of the transboundary stocks in combination with the particular characteristics of these EEZs could be influencing the results. First, both Argentine menhaden and Argentine shortfin squid have distributions with larger biomass concentration further south, with Brazil serving as the northern limit of the distributional range [@AquaMaps:2019ui; @AquaMaps:2019wt]. Second, Brazil’s EEZ is substantially larger than Uruguay’s resulting in a long distance between Brazil’s and the stocks’ centroids. Third, both stocks are expected to shift further south in the incoming years [@AquaMaps:2019ui; @AquaMaps:2019w]. Thus, the further southern shift of Argentine menhaden and shortfin squid in combination with the reduced latitudinal gradient of Uruguay’s EEZ results in a small effect in the already long latitudinal distance between Brazil’s EEZ and the stocks’ centroids. In the case of the two shrimp species, their destitution range spams along the cost of Brazil and is not expected to have a strong shift in the incoming years, also resulting in a stable transboundary index [@AquaMaps:2019wn; @AquaMaps:2019wj]. -->
